---
title: "Untitled"
author: Sophia Castellon
output: html_document
date: "2024-09-27"
---


This assignment has three questions:  

1. in a class of 35 students, Dr. J bets $1 against student's $1  that at least 2 students have the same birthday, month, day.
Find the probability that at least two student have the same birthday.  If 26 students in the class take the bet how much is Dr. J expected to win/lose.   How much should Dr. J wager against the student's $1 to make this a fair bet?
  The probability of two students have the same birthday is 0.8, Dr. J is expected is win 16.29. For it to be a fair bet Dr. J would have to wager 4.35

2. Assume class size is 26 instead of 35.  With the same bet ($1 ot $1) what is the probability Dr. J wins the bet now.  If 20 of the 26 students make the bet, what is Dr. J's expected profit/loss? 
 The probability is 0.5 and the expected profit is 4.

3. What size class (how many students) makes the $1 againt $1 a fair bet? 
  The class size that will make the $1 against $1 fair is 23. 

```{r setup, include=FALSE}
# Load necessary libraries
library(ggplot2)

# Set seed for reproducibility
set.seed(123)
```


```{r}
# Function to simulate the birthday problem
simulate_birthdays <- function(class_size = 35) {
  # Generate random birthdays for the class
  birthdays <- sample(1:365, class_size, replace = TRUE)
  
  # Check for duplicates
  any_duplicated <- length(unique(birthdays)) < class_size
  
  # Return TRUE if there are duplicates, FALSE otherwise
  return(any_duplicated)
}

# Number of simulations
n_simulations <- 100

# Vector to store probabilities
probabilities <- numeric(n_simulations)

# Run simulations
for (i in 1:n_simulations) {
  # Count how many times at least one duplicate birthday occurs
  duplicates_count <- sum(replicate(1000, simulate_birthdays()))
  
  # Calculate probability
  probabilities[i] <- duplicates_count / 1000
}

# Display probabilities
probabilities
```


```{r}
# Plot histogram of probabilities
ggplot(data.frame(probabilities), aes(x = probabilities)) +
  geom_histogram(binwidth = 0.05, fill = "lightblue", color = "black") +
  labs(title = "Histogram of Probabilities", x = "Probability", y = "Frequency") +
  theme_minimal()
```

```{r}
# Calculate mean of probabilities
mean_probability <- mean(probabilities)

# Print mean
cat("Mean = ", mean_probability, "\n")

# Calculate 95% confidence interval
ci <- quantile(probabilities, probs = c(0.025, 0.975))

# Print confidence interval
cat("95% Confidence Interval: [", ci[1], ", ", ci[2], "]\n")
```

```{r}
EV= mean_probability*26-(1-mean_probability)*26
print(paste("ev=",EV))
```

```{r}
FAIR=(mean_probability)/(1-mean_probability)
print(paste("fairbet=",FAIR))
```
## Question 2 


```{r}
# Function to simulate the birthday problem
simulate_birthdays <- function(class_size = 26) {
  # Generate random birthdays for the class
  birthdays <- sample(1:365, class_size, replace = TRUE)
  
  # Check for duplicates
  any_duplicated <- length(unique(birthdays)) < class_size
  
  # Return TRUE if there are duplicates, FALSE otherwise
  return(any_duplicated)
}

# Number of simulations
n_simulations <- 100

# Vector to store probabilities
probabilities <- numeric(n_simulations)

# Run simulations
for (i in 1:n_simulations) {
  # Count how many times at least one duplicate birthday occurs
  duplicates_count <- sum(replicate(1000, simulate_birthdays()))
  
  # Calculate probability
  probabilities[i] <- duplicates_count / 1000
}

# Display probabilities
```

```{r}
# Calculate mean of probabilities
mean_probability <- mean(probabilities)

# Print mean
cat("Mean = ", mean_probability, "\n")

# Calculate 95% confidence interval
ci <- quantile(probabilities, probs = c(0.025, 0.975))

# Print confidence interval
cat("95% Confidence Interval: [", ci[1], ", ", ci[2], "]\n")
```

```{r}
EV= mean_probability*20-(1-mean_probability)*20
print(paste("ev=",EV))
```
## Question 3

```{r}
# Function to simulate the birthday problem
simulate_birthdays <- function(class_size = 23) {
  # Generate random birthdays for the class
  birthdays <- sample(1:365, class_size, replace = TRUE)
  
  # Check for duplicates
  any_duplicated <- length(unique(birthdays)) < class_size
  
  # Return TRUE if there are duplicates, FALSE otherwise
  return(any_duplicated)
}

# Number of simulations
n_simulations <- 100

# Vector to store probabilities
probabilities <- numeric(n_simulations)

# Run simulations
for (i in 1:n_simulations) {
  # Count how many times at least one duplicate birthday occurs
  duplicates_count <- sum(replicate(1000, simulate_birthdays()))
  
  # Calculate probability
  probabilities[i] <- duplicates_count / 1000
}
mean_probability= mean(probabilities)
# Display probabilities
```

```{r}
EV= mean_probability*20-(1-mean_probability)*20
print(paste("ev=",EV))
```


